My title

### 2024-12-04 paper/paper.qmd
update the paper and the references
update the paper and the references

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Ruizi Liu
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
knitr::opts_chunk$set(
  echo = FALSE, 
  include=FALSE, 
  eval=TRUE, 
  warning = FALSE, 
  fig.width = 8,        # Set default figure width
  fig.height = 6,       # Set default figure height
  dpi = 300,            # Set default resolution for plots
  warning = FALSE,      # Suppress warnings globally
  message = FALSE       # Suppress messages globally
)

#### Loading the needed packages ####
required_packages <- c("tidyverse", "here")
for (p in required_packages) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(tidyverse)
library(here)
```

@@

#### Loading the dataset ####
data <- read_csv(here::here("data/01-raw_data/raw_data.csv"))
```

@@

@@

## Outcome variables

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

@@

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

@@

@@

### Model justification

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

@@

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: ""
subtitle: ""
author: 
  - Ruizi Liu
thanks: "Code and data are available at: [https://github.com/RIRI0527/school_shooting_analysis.git](https://github.com/RIRI0527/school_shooting_analysis.git)."
date: today
date-format: long
abstract: "School shootings remain a critical issue in the United States, with far-reaching social and psychological consequences. This study investigates the factors influencing the occurrence of school shootings, focusing on temporal trends, school type, demographic composition, and geographic context. Using a Bayesian logistic regression model, this analysis incorporates both observed and simulated data to explore the relationship between these predictors and the likelihood of a shooting. The results reveal significant associations between the predictors and school shootings. Temporal trends indicate an alarming increase in incidents over time, while public schools and urban locales exhibit higher risks compared to private schools and rural areas. Demographic composition also plays a role, with schools characterized by specific racial majorities showing distinct patterns in the likelihood of shootings. This report emphasizes the need for targeted, data-driven interventions that consider institutional and geographic variations in risk. By leveraging probabilistic modeling, this study provides actionable insights for policymakers aiming to mitigate the occurrence of school shootings and improve the safety of educational environments across the United States. "
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| include: false
#### Installing the packages ####
required_packages <- c("tidyverse", "here", "maps")
for (p in required_packages) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p, dependencies = TRUE)
    library(p, character.only = TRUE)
  }
}
```

@@

#### Loading the packages ####
library(tidyverse)
library(here)
library(maps)
library(ggplot2)
library(knitr)
library(brms)
library(bayesplot)
library(knitr)
library(kableExtra)
```

#### Loading the dataset ####
data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
shooting_data <- data %>%
  filter(data_source == "raw")
non_shooting_data <-data %>%
  filter(data_source == "synthetic")
```

# Introduction
School shootings are among the most tragic and impactful events in the United States, leaving lasting scars on communities and prompting urgent calls for effective prevention measures. While public discourse often focuses on the immediate aftermath of such events, it is equally important to examine the underlying factors contributing to their occurrence. This report aims to analyze the temporal, institutional, demographic, and geographic predictors of school shootings, drawing insights from historical data to inform future policies.

This analysis utilizes a Bayesian logistic regression model to explore whether factors such as the year, school type, demographic composition, and urbanicity influence the likelihood of school shootings. Bayesian methods are particularly suitable for this investigation, as they allow for the incorporation of prior knowledge while providing probabilistic interpretations of the results. By using both observed and simulated data, the model offers a comprehensive perspective on the risk factors associated with these incidents.

The findings presented in this report highlight critical trends and disparities that warrant further attention. For example, the temporal trend of increasing school shootings underscores the urgency of targeted interventions. Similarly, the variations in risk between public and private schools, urban and rural areas, and schools with differing demographic compositions suggest that a one-size-fits-all approach may not be sufficient. By identifying and analyzing these patterns, this study seeks to contribute to the broader conversation on preventing school shootings in the United States.

@@

The dataset is from the wasington post [@washingtonpost_school_shootings].The data provides a data record of school shootings in the United States since 1999 to 2024, with the original data set consisting of 50 variables and 416 observations, which recorded when, where, and where they occurred in the data set. The racial proportion of the schools where the shootings took place and the gender of the shooters, among other variables. These variables provide detailed information about each school shooting. We will conduct statistical analysis on this data set and explore whether these factors affect the occurrence of school shooting.

@@

Measurement data is the process by which we turn events that happen in reality into visual data by collecting information, taking measurements, and so on.
In this data, the Washington Post spent a year counting how many children were affected by the events of school shootings (not just the number of casualties).
Reporters compiled this data through Nexis, news articles, open-source databases, law enforcement reports, information from school websites, and calls to schools and police departments [@washingtonpost_school_shootings] to compile this dataset.

## Data Cleaning

We used the`R`programming language [@citeR], the `dplyr` package[@R_dplyr], the `tidyr` package [@R_tidyr], the `maps` package [@R_maps], the `ggplot2` package [@R_ggplot2], and the`scale` package [@R_scales] to clean the data.
In doing so, R code was adapted from [@tellingstories].

First we did some general cleaning of this dataset, including replacing NAs in the data with means and changing the form of dates in the data.In addition, because each school had different enrolments, we calculated the percentage of different ethnicities in each school and selected the highest percentage of ethnicities in each school.

In addition, because the original data only recorded the shootings that occurred and not the ones that did not, but we wanted to simulate the logistic regression model, we generated synthetic data based on the cleaned dataset to reflect the data that did not occur. The main components that were simulated were latitude, longitude, school ethnicity, lunch, and staffing, and to ensure that the data were reasonable, the synthetic data were generated within the allowable range of variability (a reasonable float between the maximum and minimum of the variable), and to ensure that the model was implementable, the synthetic data were generated within the allowable range of variation.Synthetic data and raw data both have 415 observations.at the same time, we made a distinction between synthetic data and raw data inside the dataset by labelling their data source in data_source and creating a new binary variable to indicate whether a shooting occurred or not.

Finally, we cleaned the dataset of some unnecessary variables, such as the name and id of the school, the specific city neighbourhoods where the shootings took place (we focused our analysis on those states specifically), the shooter's access to the gun, and so on.

In the end, we cleaned up the data to include 830 observations and 31 variables, the dataset can be find in @tbl-data-overview, only specific columns shows in the table.

## Outcome variables
The dataset reveals a notable disparity in the distribution of school types. As shown in @fig-school-type, public schools account for the vast majority of cases, while private schools are significantly underrepresented. This imbalance reflects the broader landscape of education in the United States, where public schools serve a larger student population. Moreover, this distribution is critical for our analysis, as it highlights the need to account for institutional differences when examining factors contributing to school shootings.

```{r}
#| label: fig-school-type
#| fig-cap: Distribution of School Types
#| echo: false
#| warning: false
# Count occurrences of each category in 'school_type'
school_type_counts <- as.data.frame(table(shooting_data$school_type))

# Rename the columns for clarity
colnames(school_type_counts) <- c("school_type", "Count")

# Create a bar plot
ggplot(school_type_counts, aes(x = reorder(school_type, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Distribution of School Types",
       x = "School Type",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@fig-ulocale illustrates the distribution of urban locales where school shootings occurred. Large cities and large suburbs are overrepresented compared to smaller towns and remote rural areas. This trend aligns with population density and resource disparities, suggesting that urban schools may face heightened risks due to greater exposure to social and economic stressors. Understanding the geographic distribution of incidents provides essential context for identifying at-risk locations and tailoring preventive measures accordingly.

```{r}
#| label: fig-ulocale
#| fig-cap: Distribution of Urban Locale Descriptions
#| echo: false
#| warning: false

# Assuming your dataset is loaded into a data frame called 'data'
# Count occurrences of each category in 'ulocale_desc'
ulocale_counts <- as.data.frame(table(shooting_data$ulocale_desc))

# Rename the columns for clarity
colnames(ulocale_counts) <- c("ulocale_desc", "Count")

# Create a bar plot
ggplot(ulocale_counts, aes(x = reorder(ulocale_desc, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  labs(x = "Urban Locale Description",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

The geographic distribution of school shootings across states is visualized in @fig-state-map. States such as California, Texas, and Florida exhibit higher frequencies of school shootings, reflecting their larger populations and more extensive school systems. Conversely, smaller states and those with lower populations, such as Wyoming and Vermont, report fewer incidents. This map underscores the importance of contextualizing school shootings within regional characteristics, including population size, school density, and state-specific policies.

```{r}
#| label: fig-state-map
#| fig-cap: Shooting Occurrences by State
#| echo: false
#| warning: false

# Group shooting data by state and count occurrences
state_counts <- shooting_data %>%
  group_by(state) %>%
  summarize(occurrences = n())

# Convert state names to lowercase for consistency
state_counts$state <- tolower(state_counts$state)

# Create a mapping of state abbreviations to full names
state_abbreviation_to_full <- c(
  "al" = "alabama", "ak" = "alaska", "az" = "arizona", "ar" = "arkansas", 
  "ca" = "california", "co" = "colorado", "ct" = "connecticut", "de" = "delaware", 
  "fl" = "florida", "ga" = "georgia", "hi" = "hawaii", "id" = "idaho", 
  "il" = "illinois", "in" = "indiana", "ia" = "iowa", "ks" = "kansas", 
  "ky" = "kentucky", "la" = "louisiana", "me" = "maine", "md" = "maryland", 
  "ma" = "massachusetts", "mi" = "michigan", "mn" = "minnesota", "ms" = "mississippi", 
  "mo" = "missouri", "mt" = "montana", "ne" = "nebraska", "nv" = "nevada", 
  "nh" = "new hampshire", "nj" = "new jersey", "nm" = "new mexico", "ny" = "new york", 
  "nc" = "north carolina", "nd" = "north dakota", "oh" = "ohio", "ok" = "oklahoma", 
  "or" = "oregon", "pa" = "pennsylvania", "ri" = "rhode island", "sc" = "south carolina", 
  "sd" = "south dakota", "tn" = "tennessee", "tx" = "texas", "ut" = "utah", 
  "vt" = "vermont", "va" = "virginia", "wa" = "washington", "wv" = "west virginia", 
  "wi" = "wisconsin", "wy" = "wyoming"
)

# Replace state abbreviations with full names
state_counts$state <- state_abbreviation_to_full[state_counts$state]

# Get map data for U.S. states
us_map <- map_data("state")

# Merge state data with map data
us_map_data <- left_join(us_map, state_counts, by = c("region" = "state"))

# Replace NA occurrences with 0 for states with no shootings
us_map_data$occurrences[is.na(us_map_data$occurrences)] <- 0

# Add a data frame of state centers for labeling
state_centers <- us_map %>%
  group_by(region) %>%
  summarize(
    long = mean(range(long)),
    lat = mean(range(lat))
  ) %>%
  mutate(state_abbreviation = toupper(names(state_abbreviation_to_full)[match(region, state_abbreviation_to_full)]))  # Get state abbreviations

# Plot the choropleth map with customized theme
ggplot(us_map_data, aes(x = long, y = lat, group = group, fill = occurrences)) +
  geom_polygon(color = "white") +
  geom_text(data = state_centers, aes(x = long, y = lat, label = state_abbreviation), 
            inherit.aes = FALSE, size = 3, color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "gray80") +
  theme_void() +  # Removes all axes, grid lines, and background
  theme(
    legend.position = "right",  # Keeps the legend in its default position
    plot.title = element_text(hjust = 0.5, size = 16)  # Center-align and adjust the title size
  ) +
  labs(
    fill = "Occurrences"
  )
```

The temporal trend in school shootings is depicted in @fig-year, showing a concerning increase in incidents over the years. This upward trend highlights the growing urgency of addressing the factors contributing to school shootings. Spikes in certain years may correspond to specific events or policy changes, suggesting that temporal patterns warrant further investigation to understand their underlying causes. These findings emphasize the importance of considering temporal dynamics in any analysis of school shooting data.

```{r}
#| label: fig-year
#| fig-cap: Number of Shootings by Year
#| echo: false
#| warning: false

year_counts <- as.data.frame(table(data$year))

# Rename columns for better understanding
colnames(year_counts) <- c("Year", "Count")

# Plot the data
ggplot(year_counts, aes(x = as.numeric(as.character(Year)), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
       x = "Year",
       y = "Number of Shootings") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

This analysis aimed to examine the relationship between school characteristics, demographic composition, and geographical location on the occurrence of school shootings.To achieve this, a Bayesian logistic regression model was employed. Logistic regression was deemed appropriate as the outcome variable (i.e., whether or not a school shooting occurred) is binary in nature. Bayesian methods were chosen to incorporate prior knowledge and provide probabilistic interpretations of the model parameters.

@@

The Bayesian logistic regression model can be expressed as follows:

```{=tex}
$$
y_i \mid \pi_i \sim \text{Bernoulli}(\pi_i)
$$

$$
\text{logit}(\pi_i) = \beta_0 + \beta_1 \cdot \text{year}_i + \beta_2 \cdot \text{school\_type}_i
$$
```
where:

- $y_i$: Occurrence of a school shooting (1 if a shooting occurred, 0 otherwise).

- $\pi_i$: Probability of a school shooting occurring for observation $i$.

- $\beta_0$: Intercept, representing the baseline log-odds of a school shooting for the reference categories of the predictors.

- $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$: Effects of the predictors on the log-odds of a school shooting.

**Priors**:

- $\beta_0 \sim \text{Cauchy}(0, 2)$ (Intercept prior)

- $\beta_k \sim \text{Normal}(0, 10)$ for $k = 1, 2, 3, 4$ (Predictor priors)

## Model justification

The inclusion of predictors was guided by theoretical considerations:

- **Year** ($\beta_1$): To assess temporal trends in school shootings, capturing whether incidents have increased or decreased over time.

- **School Type** ($\beta_2$): To examine if public versus private schools are differently associated with the likelihood of school shootings.

- **Dominant Race in the School Population** ($\beta_3$): To investigate the potential role of demographic composition in school shootings.

- **Urbanicity** ($\beta_4$): To explore how geographical location and urban-rural classifications affect school shooting occurrences.

Bayesian methods were used for their ability to incorporate prior knowledge and provide credible intervals for parameter estimates. Logistic regression was preferred as it ensures probabilities remain between 0 and 1, appropriate for binary outcomes.

## Diagnostics and Visualizations

To evaluate the model, several diagnostic tools and visualizations were utilized:

1.  **Posterior Predictive Check** @fig-ppc was performed to compare the observed data with predictions from the model.
2.  **Comparison of Posterior and Prior** @fig-posterior-vs-prior illustrates how the posterior distribution was updated from the priors.
3.  **Trace Plots** @fig-trace-plots were used to check for MCMC convergence.
4.  **Rhat Diagnostic Plot** @fig-rhat was used to ensure all parameters had converged ($\hat{R} < 1.1$).
5.  **90% Credibility Intervals** @fig-credible-intervals display the uncertainty around the parameter estimates.

These diagnostics confirmed good convergence and model fit, providing confidence in the estimated effects.

@@

This section reports the findings from the Bayesian logistic regression model examining the likelihood of school shootings. Parameter estimates, convergence diagnostics, and the implications of the results are presented below.

## Parameter Estimates

The parameter estimates from the Bayesian logistic regression model are shown in @fig-credible-intervals.
Positive coefficients indicate an increased likelihood of school shootings, while negative coefficients indicate a decreased likelihood.

The results highlight several significant effects:

-   **Temporal Trends**: The coefficient for `year` was positive ($\beta = 0.065$, 95% CI: \[0.042, 0.088\]), suggesting an increasing trend in school shootings over time.

-   **School Type**: Public schools were associated with higher odds of shootings compared to private schools ($\beta = 0.788$, 95% CI: \[0.069, 1.523\]).

-   **Demographic Composition**: Schools with a majority of Asian ($\beta = -6.007$, 95% CI: \[-13.346, -0.230\]) or Black students ($\beta = -5.194$, 95% CI: \[-12.128, -0.255\]) showed significantly reduced odds of school shootings.

-   **Urbanicity**: Schools in rural and suburban areas had reduced odds of shootings compared to urban schools:
    -   Remote rural areas ($\beta = -3.162$, 95% CI: \[-4.168, -2.229\])
    -   Mid-sized suburban areas ($\beta = -2.528$, 95% CI: \[-3.314, -1.792\])

These findings suggest temporal, institutional, demographic, and geographic factors all contribute to the likelihood of school shootings.

Model diagnostics confirmed the robustness of the Bayesian logistic regression.

Key diagnostic plots include:

1.  **Posterior Predictive Check** @fig-ppc: The posterior predictive distribution aligns closely with the observed data, indicating a good model fit.

2.  **Convergence**: @fig-trace-plots displays trace plots for key parameters, showing well-mixed chains that reached stationarity. All parameters had ($\hat{R} < 1.1$), as shown in @fig-rhat.

3.  **Credibility Intervals** @fig-credible-intervals: The $90%$ credibility intervals for each parameter are plotted, highlighting significant predictors (intervals that exclude 0).

These diagnostics provide strong evidence of model convergence and fit, supporting the reliability of the results.

## Implications

The results have several implications for understanding and addressing school shootings:

- The increasing trend over time underscores the urgency of targeted interventions to reduce school shootings.

- The higher risk in public schools suggests that policies and resources should prioritize these institutions.

- Geographic patterns indicate that rural and suburban schools have lower risks compared to urban schools, which may reflect differences in community or school characteristics.

- Demographic findings highlight the need for equitable and inclusive approaches to address structural inequalities potentially influencing school shootings.

Overall, these findings provide a robust framework for designing policies to mitigate the risk of school shootings.

@@

This study examined the temporal, institutional, demographic, and geographic factors influencing the likelihood of school shootings using a Bayesian logistic regression model. The model provided robust insights, with credible intervals indicating the degree of uncertainty around the estimated effects @tbl-modelresults.

## Key Findings

The results revealed several significant predictors of school shootings:

-   **Temporal Trends**: The positive association between `year` and the likelihood of school shootings highlights a concerning trend over time.
    This finding underscores the urgency of implementing effective preventative measures to reverse this trajectory.

-   **Institutional Context**: Public schools showed a significantly higher risk compared to private schools.
    This result suggests the need for targeted interventions in public school settings, potentially focusing on resource allocation, security measures, and community engagement.

-   **Demographic Composition**: The negative coefficients for schools with predominantly Asian or Black populations indicate a reduced likelihood of shootings in these contexts.

    These results call for further investigation into how demographic, cultural, or social factors may create protective environments.

-   **Geographical Variation**: Rural and suburban schools were consistently associated with lower odds of school shootings compared to urban schools.
    This pattern emphasizes the importance of location-specific prevention strategies that account for urban challenges, such as population density and resource distribution.

## Strengths and Limitations

### Strengths:

1.  **Bayesian Framework**: The use of Bayesian methods allowed for the incorporation of prior information and the quantification of uncertainty, resulting in more nuanced interpretations of the model coefficients.

2.  **Model Diagnostics**: Comprehensive diagnostic checks ([Figures 1-5](#fig-ppc)) confirmed model convergence and goodness-of-fit, lending credibility to the findings.

### Limitations:

1.  **Generalizability**: While the model captures key predictors, external factors such as policy changes or economic conditions were not included. These factors may mediate or moderate the observed relationships.

2.  **Data Availability**: The quality and completeness of the data used to fit the model may have introduced bias or omitted variable effects. Future studies should aim to include additional covariates to improve explanatory power.

## Implications and Future Directions

The findings have several implications for policy and practice:

- **Policy Interventions**: Addressing the rising trend in school shootings will require concerted efforts at the institutional level, including targeted investments in public schools.

- **Equity and Inclusion**: The protective effects associated with certain demographic groups suggest that inclusive, community-driven strategies may play a role in prevention.

- **Location-Specific Strategies**: Policymakers should tailor interventions to the unique needs of urban, suburban, and rural schools, recognizing the contextual factors that influence risk.

Future research should explore additional variables, such as socioeconomic factors, legislative impacts, and mental health resources, to provide a more comprehensive understanding of the dynamics driving school shootings.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary of the Bayesian Logistic Regression Model"

# Load the model
model <- readRDS(file = here::here("models/model_bayes.rds"))

# Extract coefficients and credible intervals
summary_model <- summary(model)
coefficients <- as.data.frame(summary_model$fixed)
coefficients <- coefficients[, c("Estimate", "l-95% CI", "u-95% CI")]

# Replace NA values with "N/A"
coefficients[is.na(coefficients)] <- "N/A"

# Format row names
rownames(coefficients) <- gsub("b_", "", rownames(coefficients))

# Create a clean table
knitr::kable(
  coefficients,
  format = "html", # Change to "latex" for PDF output or "markdown" for plain Markdown
  caption = "Summary of the Bayesian Logistic Regression Model",
  col.names = c("Estimate", "Lower CI", "Upper CI"),
  digits = 3
)
```

@@

## Appendix

### A Additional Data Details

This analysis relied on data from the Washington Post dataset, which provides detailed records of school shootings in the United States from 1999 to 2024. To enhance the robustness of the model, synthetic data were generated for schools where shootings did not occur, following the methodology described in [Section 2.3](#data-cleaning). The inclusion of synthetic data allowed for a balanced dataset of 830 observations and 31 variables.

The data cleaning process included:
- Replacing missing values with means.
- Converting date formats for consistency.
- Calculating demographic percentages for each school and retaining only the most prominent demographic group per school.
- Excluding variables that were not relevant to the analysis, such as school names and shooter-specific details.

@tbl-modelresults provides a summary of the cleaned dataset and key variables.

### B Model Details

#### B.1 Posterior Predictive Check

The posterior predictive check was performed to evaluate how well the model captures the observed data distribution. @fig-ppc demonstrates that the model predictions align closely with the observed data, indicating a good model fit.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-ppc
#| fig-cap: "Posterior predictive check"

pp_check(model, type = "dens_overlay") +
  theme_classic() +
  theme(legend.position = "bottom")
```

#### B.2 Diagnostics

To ensure the reliability of the Bayesian logistic regression model, the following diagnostics were conducted:
1. **Trace Plots:** @fig-trace-plots shows well-mixed chains for all parameters, confirming convergence.
2. **Rhat Plot:** @fig-rhat illustrates that all parameters have $\hat{R} \leq 1.1$, indicating no convergence issues.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-trace-plots
#| fig-cap: "Trace plots for selected parameters"

mcmc_trace(as.array(model), pars = c("b_year", "b_school_typepublic", "b_ulocale_descRural:Remote",
                                     "b_ulocale_descSuburb:MidMsize", "b_top_1_racesasian")) +
  theme_classic() +
  ggtitle("Trace Plots")
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-rhat
#| fig-cap: "Rhat diagnostic plot for model convergence"

# Extract Rhat values using built-in brms diagnostics
rhat_vals <- brms::rhat(model)

# Visualize Rhat values
bayesplot::mcmc_rhat(rhat_vals) +
  theme_classic() +
  ggtitle("Rhat Diagnostic Plot")
```
#### B.3 Comparison of Priors and Posteriors

@fig-posterior-vs-prior compares the posterior distributions with the priors, highlighting how the data influenced the parameter estimates.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-posterior-vs-prior
#| fig-cap: "Comparison of posterior and prior distributions"

mcmc_areas(as.array(model), pars = c("b_year", "b_school_typepublic", "b_ulocale_descRural:Remote",
                                     "b_ulocale_descSuburb:MidMsize", "b_top_1_racesasian")) +
  ggtitle("Posterior Distributions") +
  theme_classic()
```

### C Survey and Sampling Methodology

#### C.1 Population, Frame, and Sample
- **Population:** All K-12 schools in the United States.
- **Frame:** Schools from the Washington Post dataset, combined with synthetic data for schools without shootings.
- **Sample:** Stratified by school type, urban locale description, and dominant demographic group.

#### C.2 Sampling and Recruitment
- Synthetic data were generated within observed variable ranges to represent schools where shootings did not occur.
- Observed data were supplemented by demographic, geographic, and enrollment metrics.

#### C.3 Questionnaire and Validation
- **Survey Focus:** School security measures, demographic composition, geographic details, and historical incidents.
- **Validation:** Cross-referenced data with administrative records to ensure accuracy.

### D Simulation Details

To simulate the occurrence of school shootings, synthetic data were generated with random variability:
- Geographic coordinates were adjusted within observed ranges.
- Demographics and enrollment were assigned proportionally based on observed distributions.

@fig-occurrence-map visualizes the geographic distribution of shooting occurrences, highlighting both observed and synthetic data points.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-occurrence-map
#| fig-cap: School Shooting Occurrence

# Load the base map of the USA
usa_map <- map_data("state")

# Plot the map with shooting locations
ggplot() +
  geom_polygon(data = usa_map, aes(x = long, y = lat, group = group),
               fill = "gray90", color = NA) +
  geom_point(data = shooting_data, aes(x = long, y = lat),
             color = "red", alpha = 0.6, size = 2) +
  labs(
       x = "Longitude", y = "Latitude") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

### E Implications of Results

The analysis revealed the following key insights:
1. **Temporal Trends:** The increasing trend in shootings over time (@fig-trend) underscores the need for immediate interventions.

2. **Institutional and Demographic Context:** Public schools and urban locales were associated with higher risks, as shown in @fig-credible-intervals.

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-credible-intervals
#| fig-cap: "90% credibility intervals for selected predictors"

mcmc_intervals(as.array(model), pars = c("b_year", "b_school_typepublic", "b_ulocale_descRural:Remote",
                                         "b_ulocale_descSuburb:MidMsize", "b_top_1_racesasian")) +
  theme_classic() +
  ggtitle("90% Credibility Intervals")

These findings provide actionable insights for policymakers aiming to mitigate risks associated with school shootings.
